---
title: "zjin26_mini-assign4"
author: "Vincent Jin"
date: "2023-04-03"
output: pdf_document
---

# ============================================================ 
# 	Mini-assignment # 4
# ============================================================ 

# (1) calculate and store the min, max, mean, and standard deviation of all numerical columns for the following data.table:

``` {r}
library(data.table)
set.seed(100)
a = sample(65:110,1000,replace=T)
b = sample(c('M','F','F'),1000,replace=T)
c = sample(c('W','W','W','B','B','A','O',NA),1000,replace=T)
d = sample(100:5000,1000)*a
e = runif(1000,0,1)*(a/110)
dt = data.table(age = a, sex = b, race = c, cost = d, risk = e)
```

``` {r}
col <- names(dt)

for (i in col) {
  if (typeof(dt[[i]]) != "character") {
    cat(paste("the min, max, mean, sd of", i, "is :\n"))
    print(min(dt[[i]]))
    print(max(dt[[i]]))
    print(mean(dt[[i]]))
    print(sd(dt[[i]]))
  }
}
```

# (2) print mean and standard deviation (2 decimals) of the numerical columns as below (replace XXX with actual variables):
# The study population has an average age of XXX years with a standard deviation of XXX years
# The study population has an average cost of XXX dollars with a maximum of XXX dollars
# The total risk ranged between minimum of XXX and maximum of XXX

``` {r}
print(paste("The study population has an average age of", round(mean(dt$age), 2), "years with a standard deviation of", round(sd(dt$age), 2), "years"))
print(paste("The study population has an average cost of", round(mean(dt$cost), 2), "dollars with a standard deviation of", round(sd(dt$cost), 2), "dollars"))
print(paste("The total risk ranged between minimum of", round(min(dt$risk), 2), "and maximum of", round(max(dt$risk), 2)))
```

# (3) calculate the mean age, cost, and risk of each race as well as populations with missing race separately
# try using the short/alternate built-in data.table syntax (although base R syntax is also acceptable)

``` {r}
col <- c("age", "cost", "risk")
race <- unique(dt$race)
for (i in col) {
  for (r in race) {
    if (is.na(r) != TRUE) {
      print(paste("the mean", i, "of race", r, "is: "))
      print(mean(dt[[i]][which(dt$race == r)]))
    } else {
      print(paste("the mean", i, "of race", r, "is: "))
      print(mean(dt[[i]][which(is.na(dt$race))]))
    }
  }
}
```

# (4) show the histogram of age for the female population

``` {r}
hist(dt$age[which(dt$sex == "F")])
```

# (5) show the scatter plot of age vs. risk for the male population

``` {r}
plot(age ~ risk, data = dt[which(dt$sex == "M")])
```